Supplementary Information

### 2021-08-10 doc/manuscript/response_to_reviews.Rmd

Census:
I think the article mentions that most of these forest plots have repeat census data. How was this data used in the methods? If there were multiple census DBH measurements, how were these all used to constrain DBH time series?

**We have added a short paragraph to Appendix S2 specifying how DBH measurements were used: "DBH measurements were taken at the time of coring or -- for some of the trees cored within ForestGEO plots -- obtained from the most recent ForestGEO census prior to coring. For these trees, we applied the outside-in approach to DBH reconstruction (detailed below). For a minority of trees (n=50 at LDW), DBH measurements were not available, requiring use of inside-out approach to DBH reconstruction (detailed below)." We did not use multiple census measurements to constrain the DBH time series (a worthy endeavor, but beyond the scope of this project).**

Page 7-8:
RW, BAI, and deltaAGB are not independent, and BAI and delta ABH are not independent of DBH, which makes it somewhat complicated to interpret the difference in each of their relationships with climate and DBH. Further discussion on this interdependence is needed.
For example, there is reference to two relationships between RW and DBH. The first is that RW declines with age, and then second is that RW increase with DBH. Then the hypothesis that BAI peaks and then declines as a function of DBH. I don’t think it is possible (mathematically) to have RW increase with DBH be consistent with this BAI hypothesis.

**The key point with the hypotheses referenced is that we hypothesize that "[RW] *initially* increases with DBH for shade-tolerant species." This hypothesis does not preclude a subsequent decline in RW across larger size classes, which indeed was the most common outcome (described in results, "Variation with DBH" section, paragraph starting with "For RW, DBH was included..."). We have carefully reviewed the language around this issue to ensure that descriptions of an "increasing" trend in RW refer only to the smaller end of the DBH range, and have found all of the language to be accurate. While a more thorough investigation of DBH trends in growth rate and their interrelationships across growth metrics could be interesting and informative, this would be beyond the scope of the current paper.**

**More broadly, we believe that the relationships among results for the various growth metrics are fundamentally consistent. We would expect, and observe, some modest differences in climate variables selected (e.g., because the various growth metrics would effectively place different weight on above- or below-average growth rates for large vs small trees), variables included in the best model, etc. However, the results across growth metrics are broadly consistent.**

@@

 L222-224:
Dealing with tree and forest mortality is difficult. However, climate change is predicted to affect mortality in many forests. In turn, mortality can impact the growth of the remaining trees depending on the stature and location of the dead tree. This point should be addressed- the article presents the relationship between growth and climate only for the surviving trees. How does excluding growth data from the analysis potentially bias the results?

**This is a good point. Increased mortality would result in decreased competition and potential growth release for surviving individuals. Thus, in theory, we might expect increasing mortality rates in many forests around the world [@mcdowell_pervasive_2020] to result in trends of increasing growth. However, mortality trends may not be notable enough at our sites to produce this effect, or may even go in the opposite direction (e.g., reduced intermediate disturbances at CB). Moreover, increased mortality is often correlated with decreased growth unless the conditions reverse (so, for example, the trend at SC is towards increased mortality and decreased growth). Overall, in this analysis, negative growth trends predominate, and so we do not believe that growth release from increasing mortality is a dominant factor. However, we have added mention of changing mortality rates in the discussion: "Within the three older forests (ZOF, LT, CB), growth trends were mixed (Table `r hypotheses`, Fig. `r year_fig`), probably reflecting some combination of successional changes, *changing mortality rates*, and shifting competitive advantages, perhaps in part driven by changing environmental conditions [@furniss_reconciling_2017; @vrska_european_2009] or the lack of intermediate disturbances giving rise to increasing crowding [e.g., @lutz_twentiethcentury_2009]."**

### 2021-08-10 doc/manuscript/ForestGEO_treerings.Rmd
worked on reviewer comments

Growth rate -- whether measured as RW, BAI, or $\Delta$AGB -- varied with DBH for most species at all sites (Fig. `r size`).
Because the effects of calendar year could not be evaluated for all species (Fig. `r SummaryFigure`), the DBH trends described here refer to models without year. 
Relationships between growth rate and DBH had second-order terms for the majority of site-species combinations (`r round(n_spst_RW_DBH_bm/ n_spst_DBH_analyzed*100)` - `r round(n_spst_dAGB_DBH_best_model/ n_spst_DBH_analyzed*100)`% depending on growth metric; Fig. `r size`).

@@

Within the three older forests (ZOF, LT, CB), growth trends were mixed (Table `r hypotheses`, Fig. `r year_fig`), probably reflecting some combination of successional changes and shifting competitive advantages, perhaps in part driven by changing environmental conditions [@furniss_reconciling_2017; @vrska_european_2009] or the lack of intermediate disturbances giving rise to increasing crowding [e.g., @lutz_twentiethcentury_2009].
In particular, light-demanding species that establish in gaps (e.g., *Populus tremuloides* at CB; Table `r species_table`) would tend to experience an increasingly competitive environment through time.
At Zofin, size-corrected growth rates were lowest in the 1970s and 1980s, consistent with other studies from central Europe showing dramatic growth reductions due to acid deposition during this period [@elling_dendroecological_2009; @samonil_longterm_2008]. 
Non-linear trends such as this would be more accurately described by a non-linear response function to year, or incorporation of data on pollution, but that is beyond the scope of the current analysis.
Notably, there were only three species -- *Fagus sylvatica* at ZOF and *Picea pungens* and *Pinus flexilis* at CB -- whose growth rate increased significantly over the analysis time frame (Fig. `r year_fig`).

>>>>

Growth rate -- whether measured as RW, BAI, or $\Delta$AGB -- varied with DBH for most species at all sites (Fig. `r size`).
Because the effects of calendar year could not be evaluated for all species (Fig. `r SummaryFigure`), the DBH trends described here refer to models without year. 
Relationships between growth rate and DBH were best described by models with second-order terms for the majority of site-species combinations (`r round(n_spst_RW_DBH_bm/ n_spst_DBH_analyzed*100)` - `r round(n_spst_dAGB_DBH_best_model/ n_spst_DBH_analyzed*100)`% depending on growth metric; Fig. `r size`).

@@

Within the three older forests (ZOF, LT, CB), growth trends were mixed (Table `r hypotheses`, Fig. `r year_fig`), probably reflecting some combination of successional changes, changing mortality rates, and shifting competitive advantages, perhaps in part driven by changing environmental conditions [@furniss_reconciling_2017; @vrska_european_2009] or the lack of intermediate disturbances giving rise to increasing crowding [e.g., @lutz_twentiethcentury_2009].
In particular, light-demanding species that establish in gaps (e.g., *Populus tremuloides* at CB; Table `r species_table`) would tend to experience an increasingly competitive environment through time.
At Zofin, size-corrected growth rates were lowest in the 1970s and 1980s, consistent with other studies from central Europe showing dramatic growth reductions due to acid deposition during this period [@elling_dendroecological_2009; @samonil_longterm_2008]. 
Non-linear trends such as this would be more accurately described by a non-linear response function to year, or incorporation of data on pollution, but that is beyond the scope of the current analysis.
Notably, there were only three species -- *Fagus sylvatica* at ZOF and *Picea pungens* and *Pinus flexilis* at CB -- whose growth rate increased significantly over the analysis time frame (Fig. `r year_fig`).
